---
title: "SHS spatial analysis"
output:
  html_notebook: default
  pdf_document: default
---

QUESTION: What are the possible environmental drivers of SHS abundance in Apra Harbor, Guam?

AIM: Visualise publically available, spatially variable data and use PCA to identify features that characterise each sampling site.

```{r setup, include=FALSE}
current_directory <- dirname(rstudioapi::getActiveDocumentContext()$path)
knitr::opts_knit$set(root.dir=normalizePath(paste(current_directory)))
```

```{r load_libraries, warning=FALSE, message=FALSE}
## LOAD LIBRARIES ----
library(geosphere) ## distm, disGeo
library(ggbiplot) ## ggbiplot
library(viridis) ## colour blind friendly colour scheme
library(RColorBrewer) ## greys
library(ggmap) ## ggmap
library(rgdal) ## readOGR
library(raster) ## raster
library(spex) ## qm_rasterToPolygons_sp
library(ggnewscale) ## new_scale
library(ggpubr) ## ggarrange
library(vegan) ## rda
library(tidyverse)
```

```{r functions, echo=FALSE}
## ROUND2 FUNCTION ----

## create round2 function which will round up from 5 (round does not)
## https://stackoverflow.com/questions/12688717/round-up-from-5
round2 = function(x, n) {
  posneg = sign(x)
  z = abs(x)*10^n
  z = z + 0.5 + sqrt(.Machine$double.eps)
  z = trunc(z)
  z = z/10^n
  z*posneg
}

## GGBIPLOTME FUNCTION ----
# edit ggbiplot function to change arrow colour
ggbiplotme<-function (pcobj, choices = 1:2, scale = 1, pc.biplot = TRUE, 
                     obs.scale = 1 - scale, var.scale = scale, groups = NULL, 
                     ellipse = FALSE, ellipse.prob = 0.68, labels = NULL, labels.size = 3, 
                     alpha = 1, var.axes = TRUE, circle = FALSE, circle.prob = 0.69, 
                     varname.size = 3, varname.adjust = 1.5, varname.abbrev = FALSE, 
                     color = muted("red"), # <- add new arguments to the function
                     linetype = "solid", # and here
                     alpha_arrow = 1, # here too
                     ...) 
{
  library(ggplot2)
  library(plyr)
  library(scales)
  library(grid)
  stopifnot(length(choices) == 2)
  if (inherits(pcobj, "prcomp")) {
    nobs.factor <- sqrt(nrow(pcobj$x) - 1)
    d <- pcobj$sdev
    u <- sweep(pcobj$x, 2, 1/(d * nobs.factor), FUN = "*")
    v <- pcobj$rotation
  }
  else if (inherits(pcobj, "princomp")) {
    nobs.factor <- sqrt(pcobj$n.obs)
    d <- pcobj$sdev
    u <- sweep(pcobj$scores, 2, 1/(d * nobs.factor), FUN = "*")
    v <- pcobj$loadings
  }
  else if (inherits(pcobj, "PCA")) {
    nobs.factor <- sqrt(nrow(pcobj$call$X))
    d <- unlist(sqrt(pcobj$eig)[1])
    u <- sweep(pcobj$ind$coord, 2, 1/(d * nobs.factor), FUN = "*")
    v <- sweep(pcobj$var$coord, 2, sqrt(pcobj$eig[1:ncol(pcobj$var$coord), 
                                                  1]), FUN = "/")
  }
  else if (inherits(pcobj, "lda")) {
    nobs.factor <- sqrt(pcobj$N)
    d <- pcobj$svd
    u <- predict(pcobj)$x/nobs.factor
    v <- pcobj$scaling
    d.total <- sum(d^2)
  }
  else {
    stop("Expected a object of class prcomp, princomp, PCA, or lda")
  }
  choices <- pmin(choices, ncol(u))
  df.u <- as.data.frame(sweep(u[, choices], 2, d[choices]^obs.scale, 
                              FUN = "*"))
  v <- sweep(v, 2, d^var.scale, FUN = "*")
  df.v <- as.data.frame(v[, choices])
  names(df.u) <- c("xvar", "yvar")
  names(df.v) <- names(df.u)
  if (pc.biplot) {
    df.u <- df.u * nobs.factor
  }
  r <- sqrt(qchisq(circle.prob, df = 2)) * prod(colMeans(df.u^2))^(1/4)
  v.scale <- rowSums(v^2)
  df.v <- r * df.v/sqrt(max(v.scale))
  if (obs.scale == 0) {
    u.axis.labs <- paste("standardized PC", choices, sep = "")
  }
  else {
    u.axis.labs <- paste("PC", choices, sep = "")
  }
  u.axis.labs <- paste(u.axis.labs, sprintf("(%0.1f%% explained var.)", 
                                            100 * pcobj$sdev[choices]^2/sum(pcobj$sdev^2)))
  if (!is.null(labels)) {
    df.u$labels <- labels
  }
  if (!is.null(groups)) {
    df.u$groups <- groups
  }
  if (varname.abbrev) {
    df.v$varname <- abbreviate(rownames(v))
  }
  else {
    df.v$varname <- rownames(v)
  }
  df.v$angle <- with(df.v, (180/pi) * atan(yvar/xvar))
  df.v$hjust = with(df.v, (1 - varname.adjust * sign(xvar))/2)
  g <- ggplot(data = df.u, aes(x = xvar, y = yvar)) + xlab(u.axis.labs[1]) + 
    ylab(u.axis.labs[2]) + coord_equal()
  if (var.axes) {
    if (circle) {
      theta <- c(seq(-pi, pi, length = 50), seq(pi, -pi, 
                                                length = 50))
      circle <- data.frame(xvar = r * cos(theta), yvar = r * 
                             sin(theta))
      g <- g + geom_path(data = circle, color = muted("white"), 
                         size = 1/2, alpha = 1/3)
    }
    g <- g + geom_segment(data = df.v, aes(x = 0, y = 0, xend = xvar, yend = yvar), 
                          arrow = arrow(length = unit(1/2, "picas")), 
                          color = color, linetype = linetype, alpha = alpha_arrow) ## change here
  }
  if (!is.null(df.u$labels)) {
    if (!is.null(df.u$groups)) {
      g <- g + ggrepel::geom_text_repel(aes(label = labels, color = groups), ## changed here
                         size = labels.size)
    }
    else {
      g <- g + ggrepel::geom_text_repel(aes(label = labels), size = labels.size) ## changed here
    }
  }
  else {
    if (!is.null(df.u$groups)) {
      g <- g + geom_point(aes(color = groups), alpha = alpha)
    }
    else {
      g <- g + geom_point(alpha = alpha)
    }
  }
  if (!is.null(df.u$groups) && ellipse) {
    theta <- c(seq(-pi, pi, length = 50), seq(pi, -pi, length = 50))
    circle <- cbind(cos(theta), sin(theta))
    ell <- ddply(df.u, "groups", function(x) {
      if (nrow(x) <= 2) {
        return(NULL)
      }
      sigma <- var(cbind(x$xvar, x$yvar))
      mu <- c(mean(x$xvar), mean(x$yvar))
      ed <- sqrt(qchisq(ellipse.prob, df = 2))
      data.frame(sweep(circle %*% chol(sigma) * ed, 2, 
                       mu, FUN = "+"), groups = x$groups[1])
    })
    names(ell)[1:2] <- c("xvar", "yvar")
    g <- g + geom_path(data = ell, aes(color = groups, group = groups))
  }
  if (var.axes) {
    g <- g + geom_text(data = df.v, aes(label = varname, 
                                        x = xvar, y = yvar, angle = angle, hjust = hjust), 
                       color = color, size = varname.size) ## changed here too
  }
  return(g)
}

```

```{r PCA_data, warning=FALSE, message=FALSE}
## DEPTH ----

## import field sample data (including eDNA concentration)
SHS_field_sample_data_unmod <- read.csv("docs/00.2_field_sample_data.csv")

## retrieve site depth info
depth_data <- SHS_field_sample_data_unmod %>%
  ## extract information for first field sample each site/month
  ## (depth was most often recorded once per day per site)
  filter(., sample_no == 1) %>%
  ## convert depth to m
  mutate(depth_m = depth_ft*0.3048) %>%
  ## select relevant columns 
  select(., site_name, latitude, longitude, depth_m) %>%
  ## format lat lon
  mutate(latitude = round2(as.numeric(latitude), 5), 
         longitude = round2(as.numeric(longitude), 4)) %>%
  ## remove na values
  drop_na()

## TSS ----

## import sentinel 2 formatted data 
S2_site <- read.csv("docs/sentinel2_site_level_data.csv")

## retrieve site TSS info
TSS_data <- S2_site %>%
  ## format lat lon
  mutate(latitude = round2(as.numeric(lat), 5), 
         longitude = round2(as.numeric(lon), 4)) %>%
  ## calculate average TSS for each site number
  group_by(site_name, site_num, latitude, longitude) %>%
  summarise_at(vars(mean_green_red = green_red), mean)

## COMBINE AND CALCULATE DISTANCE ----
depth_TSS_dist_data <- depth_data %>%
  ## combine depth and TSS by GPS points 
  left_join(., TSS_data, by=c("latitude", "longitude", "site_name")) %>%
  ## calculate distance from each point to harbor entrance:
  mutate(entrance_lat = 13.4512, entrance_lon = 144.6179) %>%
  ## calculate the distance between the two sets of points 
  rowwise() %>%
  mutate(distance = distm(x=c(entrance_lon, entrance_lat), 
                          y=c(longitude, latitude), 
                          fun = distGeo)) %>%
  ## Blue Hole is outside the harbor, so multiply these values by -1
  mutate(distance = ifelse(site_name == "Blue Hole", distance*-1, distance)) %>%
  ## format lat lon
  mutate(latitude = round2(as.numeric(latitude), 4), 
         longitude = round2(as.numeric(longitude), 4))

## BENTHIC DATA ---- 

## import benthic data
benthic_data_unmod <- read.csv("docs/guam_first_points_intersection.csv")

## format data
spatial_data <- benthic_data_unmod %>%
  ## rename
  rename(latitude = Latitude, longitude = Longitude) %>%
  ## merge data sets (site numbers required)
  left_join(., depth_TSS_dist_data, 
            by=c("latitude", "longitude")) %>%
  ## rename columns (from metadata file provided by NCCOS)
  rename(major_reef_structure = M_STRUCT,
         geomorphologic_zone = ZONE,
         major_biological_cover = M_COVER) %>%
  ## remove land info (due to buffer overlap)
  filter(., geomorphologic_zone != "Land") %>%
  ## calculate total area per site_num
  group_by(site_num) %>%
  add_tally(., wt=intersection_area, name = "total_area") %>%
  ## calculate % area
  mutate(percent_area = intersection_area/total_area*100) 

spatial_data_mbc <- spatial_data %>%
  ## calculate the total area occupied by major_biological_cover
  group_by(site_num, major_biological_cover) %>%
  add_tally(., wt=percent_area, name = "percent_area_sum") %>%
  ## select columns and remove duplicate rows
  ungroup() %>%
  select(site_name, site_num, depth_m, mean_green_red, 
         distance, major_biological_cover, percent_area_sum, 
         latitude, longitude) %>%
  unique(.) %>%
  ## format wide for pca
  spread(., key=major_biological_cover, value=percent_area_sum) %>%
  ## replace missing values with zeros
  replace(., is.na(.), 0) %>%
  ## remove unknown
  select(-Unknown)

spatial_data_mrs <- spatial_data %>%
  ## calculate the total area occupied by major_reef_structure
  group_by(site_num, major_reef_structure) %>%
  add_tally(., wt=percent_area, name = "percent_area_sum") %>%
  ## select columns and remove duplicate rows
  ungroup() %>%
  select(site_name, site_num, depth_m, mean_green_red, 
         distance, major_reef_structure, percent_area_sum, 
         latitude, longitude) %>%
  unique(.) %>%
  ## format wide for pca
  spread(., key=major_reef_structure, value=percent_area_sum) %>%
  ## replace missing values with zeros
  replace(., is.na(.), 0) %>%
  ## remove unknown
  select(-Unknown)

spatial_data_full <- spatial_data_mbc %>%
  left_join(., spatial_data_mrs, 
            by = c("site_name", "site_num", "depth_m", "mean_green_red", "distance", "latitude", "longitude"))

## export it
write.csv(spatial_data_full, "docs/04.0_spatial_data_full.csv", row.names=FALSE)
```

```{r sum_stats, warning=FALSE, message=FALSE, fig.width=7, fig.height=7}
## create table and plot of summary statistics ----

spatial_data_sum <- spatial_data_full %>%
   select(site_name, depth_m, mean_green_red, distance, Coral, Macroalgae, Turf, 
          Uncolonized, 'Coral Reef and Hardbottom', 'Unconsolidated Sediment') %>%
   ## add nicer names
   rename('Depth (m)' = 'depth_m',
              'Turbidity (B3:B4)' = 'mean_green_red',
              'Distance from entrance (m)' = 'distance',
              '% Coral cover' = 'Coral',
              '% Macroalgal cover' = 'Macroalgae',
              '% Turf algal cover' = 'Turf',
              '% Uncolonised' = 'Uncolonized', 
              '% Coral reef and hardbottom' = 'Coral Reef and Hardbottom',
              '% Unconsolidated sediment' = 'Unconsolidated Sediment') %>%
   group_by(site_name) %>%
   summarise(across(everything(), list(mean = mean, SD = sd))) %>%
   rename('Site' = site_name)

colnames(spatial_data_sum) <- gsub("_", " ", colnames(spatial_data_sum))

write.csv(spatial_data_sum, "docs/04.1_spatial_data_summary.csv", row.names = FALSE)

spatial_data_long_less <- spatial_data_full %>%
   select(site_name, depth_m, mean_green_red, distance, Coral, Macroalgae, Turf, Uncolonized, 'Coral Reef and Hardbottom', 'Unconsolidated Sediment') %>%
   gather(., -site_name, key = "variable", value = "value") %>%
   ## add nicer names
   mutate(variable_name = gsub('depth_m', 'Depth (m)', variable) %>%
                         gsub('mean_green_red', 'Turbidity (B3:B4)', .) %>%
                         gsub('distance', 'Distance from entrance (m)', .) %>%
                         gsub('Coral', '% Coral cover', .) %>%
                         gsub('Macroalgae', '% Macroalgal cover', .) %>%
                         gsub('Turf', '% Turf algal cover', .) %>%
                         gsub('Uncolonized', '% Uncolonised', .) %>%
                         gsub('% Coral cover Reef and Hardbottom', '% Coral reef and hardbottom', .) %>%
                         gsub('Unconsolidated Sediment', '% Unconsolidated sediment', .))

spatial_data_plot <- ggplot(spatial_data_long_less, aes(x = site_name, y = value, fill = site_name))+
   geom_boxplot()+
   facet_wrap(facets = "variable_name", scales = "free")+
   theme_bw()+
   labs(x = "Site", y = "Value")+
   theme(axis.text.x = element_text(angle = 45, hjust = 1),
                   panel.grid.minor = element_blank(), 
                   legend.position="none")+
   scale_fill_viridis(discrete = TRUE)

spatial_data_plot

ggsave("docs/figure_supp_spatial_data_box.tiff", spatial_data_plot, device="tiff", 
               width = 210, height = 297/1.5, units="mm")

```

```{r PCA, fig.width=3.5, fig.height=3, warning=FALSE, message=FALSE}
## PCA ----

## edit column names for plot
pca_data_full <- spatial_data_full %>%
  select(-latitude, -longitude) %>%
  rename(., "Distance to entrance" = distance) %>%
  rename(., Turbidity = mean_green_red) %>%
  rename(., Depth = depth_m) %>%
  rename(., "Structure: Sediment\n" = "Unconsolidated Sediment") %>%
  rename(., "Structure: Reef and hardbottom" = "Coral Reef and Hardbottom") %>%
  rename(., "Cover: Coral" = Coral) %>%
  rename(., "Cover: Macroalgae" = Macroalgae) %>%
  rename(., "Cover: Turf algae" = Turf) %>%
  rename(., "\nCover: Uncolonised" = Uncolonized)

## compute
pca_full<-prcomp(pca_data_full[,c(3:ncol(pca_data_full))], center=TRUE, scale. = TRUE)

pca_plot<-ggbiplotme(pca_full, 
                      varname.size = 3,
         groups=pca_data_full$site_name,
         color="black",
         alpha=0,
         ellipse.alpha=0.1,
         ellipse=TRUE
         )+
  # coord_cartesian(xlim = c(-3.9, 3.9), ylim=c(-3.1, 2.2)) +
  coord_cartesian(xlim = c(-3.9, 3.9), ylim=c(-2.2, 3.2)) +
  xlab("PC1 (54.3 % explained variance)")+
  ylab("PC2 (16.3 % explained variance)")+
  scale_colour_viridis(discrete="TRUE", 
                       labels=c("Blue\nHole     ", "Inner\nHarbor     ", 
                                "Middle\nShoals     ", "Orote\nPoint     ", "Sasa\nBay     "))+
  geom_point(aes(colour=pca_data_full$site_name), size=5, alpha=0.4)+
  theme_bw()+
  theme(legend.position = "bottom", 
        plot.margin=unit(c(0.5,0.2,0.75,0.4),"cm"))+ #TRBL
  guides(color=guide_legend("Site"))

# pca_plot
```

```{r RDA, fig.width=7, fig.height=6, warning=FALSE, message=FALSE}
## z-score transform data (due to different base units of measurement)

# rda_data <- pca_data_full %>%
#   mutate(sample = paste(site_name, row_number(), sep=" ")) %>%
#   column_to_rownames(., var="sample") %>%
#   mutate(across(Depth:'Structure: Sediment\n', ~scale(.)))
# 
# rda_sites <- pca_data_full %>%
#   select(site_name)
# 
# ## check number of species columns 
# names(rda_data)
# rda_anal <- rda(rda_data[,c(3:ncol(rda_data))]~site_name, data=rda_data, scale=T)   
# 
# ## check summary
# summary(rda_anal)
# 
# ## check for significance and amoung of vairance explained by each term
# rda_aov<-anova.cca(rda_anal, by = "terms")
# rda_aov
```

```{r benthic_map_data, warning=FALSE, message=FALSE}
## MAKE A GOOGLE MAP OF APRA HARBOR ----

## provide ggmap with your key
# register_google(key = "[your key here]") 

## create map style at https://mapstyle.withgoogle.com/
s2<- "style=element:geometry%7Ccolor:0xf5f5f5&style=element:labels%7Cvisibility:off&style=element:labels.icon%7Cvisibility:off&style=element:labels.text.fill%7Ccolor:0x616161&style=element:labels.text.stroke%7Ccolor:0xf5f5f5&style=feature:administrative%7Celement:geometry%7Cvisibility:off&style=feature:administrative.land_parcel%7Cvisibility:off&style=feature:administrative.land_parcel%7Celement:labels.text.fill%7Ccolor:0xbdbdbd&style=feature:administrative.neighborhood%7Cvisibility:off&style=feature:poi%7Cvisibility:off&style=feature:poi%7Celement:geometry%7Ccolor:0xeeeeee&style=feature:poi%7Celement:labels.text.fill%7Ccolor:0x757575&style=feature:poi.park%7Celement:geometry%7Ccolor:0xe5e5e5&style=feature:poi.park%7Celement:labels.text.fill%7Ccolor:0x9e9e9e&style=feature:road%7Cvisibility:off&style=feature:road%7Celement:geometry%7Ccolor:0xffffff&style=feature:road%7Celement:labels.icon%7Cvisibility:off&style=feature:road.arterial%7Celement:labels.text.fill%7Ccolor:0x757575&style=feature:road.highway%7Celement:geometry%7Ccolor:0xdadada&style=feature:road.highway%7Celement:labels.text.fill%7Ccolor:0x616161&style=feature:road.local%7Celement:labels.text.fill%7Ccolor:0x9e9e9e&style=feature:transit%7Cvisibility:off&style=feature:transit.line%7Celement:geometry%7Ccolor:0xe5e5e5&style=feature:transit.station%7Celement:geometry%7Ccolor:0xeeeeee&style=feature:water%7Celement:geometry%7Ccolor:0xc9c9c9&style=feature:water%7Celement:labels.text.fill%7Ccolor:0x9e9e9e&size=480x360"
## (etc.)
## get a Google map of Apra Harbor (retrieves raster map from Google Maps)
## Note that in most cases by using this function you are agreeing to 
## the Google Maps API Terms of Service at:
## https://cloud.google.com/maps-platform/terms/ **

## (won't work without key)
ApraHarbor <- get_googlemap(center = c(lon = 144.6521, lat = 13.44297),
                            zoom = 13,
                            maptype = 'terrain',
                            color = 'bw',
                            style = s2)

## use ggmap to plot the the raster using the ggplot2 framework
# ggmap(ApraHarbor)

## MAKE HABITAT AND STRUCTURE MAPS OF APRA HARBOR ----

## read shape file and layer into a spaital vector object
guam_habitat <- readOGR("./docs", "guam_habitat", verbose=FALSE)

## re-project data
guam_habitat <- spTransform(guam_habitat, CRS("+proj=longlat +datum=WGS84"))

## fortify data for ggplot2
guam_habitat_fortify <- fortify(guam_habitat)

## add and id field for geom_polygon
guam_habitat@data$id <- 0:(dim(guam_habitat@data)[1]-1) 

## join fortified df and habitat data by id
guam_habitat_join = plyr::join(x = guam_habitat_fortify, 
                               y = guam_habitat@data, 
                               by="id") # join by id

## remove unhelpful polygons for structure category 
guam_habitat_structure <- guam_habitat_join %>%
  filter(., M_STRUCT != "Unknown") %>%
  filter(., M_STRUCT != "Other Delineations")

## remove unhelpful polygons for cover category 
guam_habitat_cover <- guam_habitat_join %>%
  filter(., M_COVER != "Unclassified") %>%
  filter(., M_COVER != "Unknown") %>%
  filter(., M_COVER != "Seagrass") # %>% ## (none recorded)
  # filter(., M_COVER != "Coralline Algae") ## (none recorded)
 
## create some colour scales 
five_greys <- brewer.pal(6, "Greys")[2:6]
two_greys <- brewer.pal(3, "Greys")[2:3]
```

```{r plot_benthic_cover, warning=FALSE, message=FALSE}
## PLOT COVER ----

## create sampling points df
sampling_points <- benthic_data_unmod %>%
  select(Site.Name, Longitude, Latitude) %>%
  unique(.)

## edit categories names to correspond to PCA
guam_habitat_cover_mod <- guam_habitat_cover %>%
  mutate(M_COVER = gsub("Emergent Vegetation", "Vegetation", M_COVER),
         M_COVER = gsub("Turf", "Turf algae", M_COVER),
         M_COVER = gsub("Uncolonized", "Uncolonised", M_COVER), 
         ## re-order for nice legend
         M_COVER = factor(M_COVER, levels=c("Coralline Algae", "Coral", 
                                            "Macroalgae", "Turf algae", 
                                            "Uncolonised", "Vegetation")), 
         ## re-order so middle shoals polygons are visible
         group2 = fct_rev(group))

## create plot
cover_map <- ggmap(ApraHarbor) + 
  geom_polygon(data=guam_habitat_cover_mod, 
               aes(x=long, y=lat, group=group2, 
                   fill=M_COVER))+
  scale_fill_manual(values = brewer.pal(6, "Greys")) +
  labs(fill = "Major\nbiological\ncover") +
  # limit the size of the plot (produces message, ignore it)
  scale_x_continuous(limits = c(144.615, 144.688), expand = c(0, 0)) +
  scale_y_continuous(limits = c(13.415, 13.4675), expand = c(0, 0)) +
  new_scale("fill")+
  geom_point(data = sampling_points, 
             aes(x=Longitude, y=Latitude, 
                 colour=Site.Name, 
                 fill=Site.Name), 
             size=4.5, shape=21) +
  scale_fill_viridis(discrete=TRUE, alpha=0.3) +
  scale_colour_viridis(discrete=TRUE)+
  guides(fill=FALSE, colour=FALSE)+
  theme(legend.position = "bottom", 
        panel.border = element_rect(colour = "black", fill=NA, size=0.5))

# cover_map
```

``` {r plot_benthic_structure, warning=FALSE, message=FALSE}
## PLOT STRUCTURE ----

## edit categories names to correspond to PCA
guam_habitat_structure_mod <- guam_habitat_structure %>%
  mutate(M_STRUCT = gsub("Coral Reef and Hardbottom", "Reef and Hardbottom", M_STRUCT),
         M_STRUCT = gsub("Unconsolidated Sediment", "Sediment", M_STRUCT),
         ## relevel so middle shoals coral polygons are visible
         group2 = fct_rev(group))

## create plot
structure_map <- ggmap(ApraHarbor) + 
  geom_polygon(data=guam_habitat_structure_mod, 
               aes(x=long, y=lat, group=group2, 
                   fill=M_STRUCT))+
  scale_fill_manual(values = two_greys) +
  labs(fill = "Major\nreef\nstructure") +
  # limit the size of the plot (produces message, ignore it)
  scale_x_continuous(limits = c(144.615, 144.688), expand = c(0, 0)) +
  scale_y_continuous(limits = c(13.415, 13.4675), expand = c(0, 0)) +
  guides(fill=FALSE, colour=FALSE)+
  new_scale("fill")+
  geom_point(data = sampling_points, 
             aes(x=Longitude, y=Latitude, 
                 colour=Site.Name, 
                 fill=Site.Name), 
             size=4.5, shape=21) +
  scale_fill_viridis(discrete=TRUE, alpha=0.3) +
  scale_colour_viridis(discrete=TRUE) +
  theme(legend.position = "bottom", 
        panel.border = element_rect(colour = "black", fill=NA, size=0.5))

# structure_map
```

```{r depth_data, warning = FALSE, message=FALSE}
## IMPORT AND FORMAT DEPTH DATA FOR APRA HARBOR ----
## NB: This chunk will take a long time to run

## data source:
## https://pae-paha.pacioos.hawaii.edu/thredds/ncss/ngdc_bathy_10m_guam/dataset.html
## coords: 13.4-13.5, 144.6-144.7 

## try 10m data, cropped for region of interest
depth_10m_guam <- raster("docs/ngdc_bathy_10m_guam_cropped.nc")

## sp returns SpatialPolygonsDataFrame 
depth_10m_guam_qmrtpsp <- qm_rasterToPolygons_sp(depth_10m_guam, na.rm=TRUE)

## add id field (for geom_polygon)
depth_10m_guam_qmrtpsp@data$id <- 0:(dim(depth_10m_guam_qmrtpsp@data)[1]-1) 

## fortify (for ggplot2)
depth_10m_guam_qmrtpsp_fortify <- fortify(depth_10m_guam_qmrtpsp) 

## join now dataframes by id (for geom_polygon)
depth_10m_join <- plyr::join(x = depth_10m_guam_qmrtpsp_fortify, 
                               y = depth_10m_guam_qmrtpsp@data, 
                               by="id") 

## remove deep points (so shading is more informative)
## - 200 based on 1m Apra map at max depth 186 m
depth_10m_join_less <- filter(depth_10m_join, elevation > -200) 
```

```{r plot_depth, warning=FALSE, message=FALSE}
## PLOT DEPTH ----

## map depths less than 200 m deep, with sampling points
depth_map <- ggmap(ApraHarbor) + 
  geom_polygon(data=depth_10m_join_less, 
               aes(x=long, y=lat, group=group, 
                   fill=elevation*-1)) +
  scale_x_continuous(limits = c(144.615, 144.688), expand = c(0, 0)) +
  scale_y_continuous(limits = c(13.415, 13.4675), expand = c(0, 0)) +
  scale_fill_gradient(low="gray90", high="gray0") +
  labs(fill = "Depth\n(m)") +
  guides(group=FALSE, fill=FALSE, colour=FALSE) +
  new_scale("fill")+
  geom_point(data = sampling_points, 
             aes(x=Longitude, y=Latitude, 
                 colour=Site.Name, 
                 fill=Site.Name), 
             size=4.5, shape=21) +
  scale_fill_viridis(discrete=TRUE, alpha=0.3) +
  scale_colour_viridis(discrete=TRUE) +
  theme(legend.position = "bottom", 
        panel.border = element_rect(colour = "black", 
                                    fill=NA, size=0.5))

# depth_map
```

```{r combine_plots, warning=FALSE, message=FALSE, fig.height=9.4, fig.width=9.9}
## COMBINE PLOTS ----
figure_spatial<-ggarrange(structure_map,
                 cover_map,
                 depth_map, 
                 pca_plot,
                 nrow = 2, ncol = 2,
                 labels = c("A", "B", "C", "D"))
figure_spatial
```

#### Result: 

Based on overly simplistic but full-coverage PacIOOS data for major reef structure, major biological cover and depth in Apra Harbor, as well as senitnel 2 data for turbidity and the distance of each sampling point to the harbor entrance, sampling points in our key SHS sites (Sasa Bay; yellow, Inner Harbor; blue) are characterised by sediment as the major reef structure, uncolonised biological cover or (for sasa bay) macroalgae, high turbidity, shallow depths and large distances from the harbor entrance.

```{r save_plot, include=FALSE}
ggsave("docs/figure_spatial_2.tiff", figure_spatial, device="tiff", 
       width = 210*1.2, height = 297*0.8, units="mm", dpi = 500)
```

